\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning on Unsecured Trajectories

### 2023-10-23 paper.tex
update

This section presents the two components of our approach: constructing the checkpoint graph based on unsecured multi-robot trajectories, and the formulation and solution of the network flow problem.

@@

\subsection{Co-observation planning problem}
In this section, we describe in detail how to formulate the cross-trajectory planning problem as a network flow problem, and solve it using mixed-integer linear programs (MILP). We assume that $n_p$ \emph{reference robots} are dedicated (one in each sub-team) to follow the reference trajectory.  The goal of this section is to minimize the number and plan the routes of $n-n_p$  additional \emph{cross-trajectory robots} dedicated to perform co-observations with the reference robots.

\begin{remark}
Note that we assign fixed roles to reference robots for convenience in explaining the multiflow formulation. In practice, after a cross-trajectory robot joins a team, it is considered interchangable, and could switch roles with the reference robot of that trajectory (i.e., the original reference robot might serve as a cross-trajectory robot in a later cross-trajectory co-observation).
\end{remark}

@@

Our goal is to find a path in $G$ for all the cross-trajectory robots that starts from the virtual source $v^{+}$ and ends in the virtual sink $v^{-}$.
The path of a robot $k$ can be represented as a flow vector $\vf^{k} = \{ f^{k}_{ij} \}$, where $f^{k}_{ij} \in \{1,0\}$ is an indicator variable representing whether robot $k$'s path contains the edge $v_{i}\to v_{j}$. 
As required by the security guarantee in \Cref{rmk:revised-security}, reference robots must be seen by at least one cross-reference robot at every checkpoint.
Thus, the planning problem can then be formulated as a path cover problem on $G_{q}$, i.e., as finding a set of paths $F=[\vf_{1},\dots, \vf_{\cK}]$ for cross-trajectory robots such that every checkpoint in $V_{p}$ is included in at least one path in $F$. Additionally, to encourage the robots' exchange between different sub-teams, cross-trajectory co-observations should be preferred compared to co-observation within the same team. \rtron{Finally, edges from the virtual source and to the virtual sink should have zero cost, to allow robots to automatically get assigned to the starting point that is most convenient for the overall solution.}
These requirements are achieved with the weights for edges $(v_{i},v_{j})\in E$ defined as:
\begin{equation}
	w_{i,j}=\begin{cases}
	-w_{t} & \cI_{v_{i}}=\cI_{v_{j}}, (v_{i},v_{j})\in E_{q}\\
	w_{c} & \cI_{v_{i}} \neq \cI_{v_{j}}, (v_{i},v_{j})\in E_{q}\\
	0 &  (v_{i},v_{j})\in E / E_{q} 
	\end{cases}
\end{equation}
where $w_{c} > w_{t}$.

With the formulation above, the planning problem can be written as an optimization problem, where the optimization cost balances between the co-observation performance and total number of flows (cross-trajectory robots) needed:
 \begin{subequations} \label{eq:flow-coverage-problem}
     \begin{align}
        \min & \sum^{\cK}_{k} \sum_{(+i)\in E} f^{k}_{+i} - \rho \sum^{\cK}_k \sum_{(ij)\in E} w_{ij} f^k_{ij} \label{eq:flow-cost}\\
        s.t. & \sum_{\{h:(hi) \in E\}}f^k_{hi}=\sum_{\{j:(ij) \in E\}}f^k_{ij},  \forall k,\forall v \in V_{q} / \{v^{+}, v^{-}\}  \label{eq:FlowBalanceConstraint}\\
        & \sum^{\cK}_k\sum_{\{i:(ij)\in E \}} f^k_{ij} \geq 1, \forall v_{j} \in \{V^{s}_{p}\} \label{eq:FlowCoverage}\\
        & f^k_{ij} \in \{0,1\} ,  \forall (ij)\in E\label{eq:SingleFLowCapacity}
     \end{align}
 \end{subequations}
where, for convenience, we used $(ij)$ to represent the edge $(v_{i},v_{j})$, and $(+i)$ to represent the edge $(v^{+},v_{i})$.

The first term in \ref{eq:flow-cost} is total outgoing flow from the source $v^{+}$ while the second term is the total cost of all flows representing the overall co-observation performance (defined as the total number of cross-trajectory edges taken by all flows beyond the regular trajectory edges). The constant $\rho$ is a manually selected penalty parameter to balance between two terms in the cost function. The constraint \eqref{eq:FlowBalanceConstraint} is the flow conservation constraint, which ensures that the amount of flow entering and leaving a given node $v$ is equal (except for $v^{+}$ and $v^{-}$). The constratin \eqref{eq:FlowCoverage} is the flow coverage constraints, which ensures that all security checkpoints $ \{V^{s}_{p}\}$ have been visited by at least one flow (robot). Since the security graph $G_{q}$ is acyclic, this problem is in complexity class $P$, and can be solved in polynominal time \cite{ntafos1979path}.

\subsection{Co-observation performance}
Notice that problem \eqref{eq:flow-coverage-problem} is guaranteed to have a solution for $\cK=N_{p}$ where all $\cK$ robots follows the reference trajectory ($f^{k}_{ij}=1, \forall \cI_{v_{i}}=\cI_{v_{j}}=k$).  Additionally, for a fixed number of robots $\cK \geq N_{p}$, it is possible to have a subset of flows $F_{\cE} \in F$ that is empty, i.e. $f_{ij}=0, \forall (v_{i},v_{j})\in E$, $f\in F_{\cE}$, these flow will not increase the cost and will not have any impact on the solution.  
The trade-off between the number of surveillance robots used, and the overall security performance is a critical consideration in this case. Increasing the number of robots generally contributes to improved security by enabling more cross-trajectory co-observations. However, adding more robots can lead to diminishing returns, as the benefits gained from additional robots may be counterbalanced by increased complexity and coordination challenges. To capture this trade-off, the penalty parameter $\rho$ in the cost function \ref{eq:flow-cost} allows us to fine-tune the balance between the number of robots employed and the desired task performance. By optimizing this parameter, we can strike a suitable equilibrium between security enhancement and operational efficiency.

To find such tradeoff, we employ an iterative approach. We start with $\cK=1$ and gradually increase it until $F$ contains a empty flow. This indicates that the performance can no longer be improved by adding more robots. To ensure that the value of $\cK$ does not grow unbounded, penalty parameter $\rho$ need to ensure that the second term for a single flow in \eqref{eq:flow-cost} $ \rho \sum_{(ij)\in E} w_{ij} f^k_{ij}$ is always smaller than $1$.

@@

\section{Summary}
In this paper, we provide a method to enhance the security of a multi-robot system without sacrificing the performance. This is done by introducing additional robots to perform cross-trajectory co-observations by traveling between different trajectories. We model the unsecured multi-robot trajectories as a checkpoint graph by identifying checkpoints that requires observation and cross-trajectory paths that can safely access the checkpoints from different trajectories. We have shown that the co-observation planning problem across different trajectories can be transformed into a multi flow problem on the graph, and that we can find the minimal number of robot needed to finish the co-observation task.

>>>>

This section presents the two components of our approach: constructing the checkpoint graph based on unsecured multi-robot trajectories, and the formulation and solution of the network multi-flow problem.

@@

\subsection{Co-observation planning problem}
In this section, we describe in detail how to formulate the cross-trajectory planning problem as a network multi-flow problem, and solve it using mixed-integer linear programs (MILP). We assume that $n_p$ \emph{reference robots} are dedicated (one in each sub-team) to follow the reference trajectory.  The goal of this section is to minimize the number and plan the routes of $n-n_p$  additional \emph{cross-trajectory robots} dedicated to perform co-observations with the reference robots.

\begin{remark}
Note that we assign fixed roles to reference robots for convenience in explaining the multi-flow formulation. In practice, after a cross-trajectory robot joins a team, it is considered interchangable, and could switch roles with the reference robot of that trajectory (i.e., the original reference robot might serve as a cross-trajectory robot in a later cross-trajectory co-observation).
\end{remark}

@@

Our goal is to find a path in $G$ for all the cross-trajectory robots that starts from the virtual source $v^{+}$ and ends in the virtual sink $v^{-}$.
The path of a robot $k$ can be represented as a flow vector $\vf^{k} = \{ f^{k}_{ij} \}$, where $f^{k}_{ij} \in \{1,0\}$ is an indicator variable representing whether robot $k$'s path contains the edge $v_{i}\to v_{j}$. 
In order to perform co-observations as required by the security introduced in \Cref{rmk:revised-security}, reference robots must be seen by at least one cross-trajectory robot at every checkpoint.
Thus, the planning problem can be formulated as a path cover problem on $G_{q}$, i.e., as finding a set of paths $F=[\vf_{1},\dots, \vf_{\cK}]$ for cross-trajectory robots such that every checkpoint in $V_{p}$ is included in at least one path in $F$. Additionally, to encourage the robots' exchange between different sub-teams, cross-trajectory co-observations should be preferred compared to co-observation within the same team.

Finally, edges from the virtual source and to the virtual sink should have zero cost, to allow robots to automatically get assigned to the starting point that is most convenient for the overall solution.
These requirements are achieved with the weights for edges $(v_{i},v_{j})\in E$ defined as:
\begin{equation}
	w_{i,j}=\begin{cases}
	-w_{t} & \cI_{v_{i}}=\cI_{v_{j}}, (v_{i},v_{j})\in E_{q}\\
	w_{c} & \cI_{v_{i}} \neq \cI_{v_{j}}, (v_{i},v_{j})\in E_{q}\\
	0 &  (v_{i},v_{j})\in E / E_{q} 
	\end{cases}
\end{equation}
where $w_{c} > w_{t}$.

With the formulation above, the planning problem can be written as an optimization problem, where the optimization cost balances between the co-observation performance and total number of flows (cross-trajectory robots) needed:
 \begin{subequations} \label{eq:flow-coverage-problem}
     \begin{align}
        \min_{F} & \sum^{\cK}_{k} \sum_{(+i)\in E} f^{k}_{+i} - \rho \sum^{\cK}_k \sum_{(ij)\in E} w_{ij} f^k_{ij} \label{eq:flow-cost}\\
        s.t. & \sum_{\{h:(hi) \in E\}}f^k_{hi}=\sum_{\{j:(ij) \in E\}}f^k_{ij},  \forall k,\forall v \in V_{q} / \{v^{+}, v^{-}\}  \label{eq:FlowBalanceConstraint}\\
        & \sum^{\cK}_k\sum_{\{i:(ij)\in E \}} f^k_{ij} \geq 1, \forall v_{j} \in \{V^{s}_{p}\} \label{eq:FlowCoverage}\\
        & f^k_{ij} \in \{0,1\} ,  \forall (ij)\in E\label{eq:SingleFLowCapacity}
     \end{align}
 \end{subequations}
where, for convenience, we used $(ij)$ to represent the edge $(v_{i},v_{j})$, and $(+i)$ to represent the edge $(v^{+},v_{i})$.

The first term in \ref{eq:flow-cost} is total outgoing flow from the source $v^{+}$ while the second term is the total cost of all flows representing the overall co-observation performance (defined as the total number of cross-trajectory edges taken by all flows beyond the regular trajectory edges). The constant $\rho$ is a manually selected penalty parameter to balance between two terms in the cost function. The constraint \eqref{eq:FlowBalanceConstraint} is the flow conservation constraint, which ensures that the amount of flow entering and leaving a given node $v$ is equal (except for $v^{+}$ and $v^{-}$). The constratin \eqref{eq:FlowCoverage} is the flow coverage constraints, which ensures that all security checkpoints $ \{V^{s}_{p}\}$ have been visited by at least one flow (robot). Since the security graph $G_{q}$ is acyclic, this problem is in complexity class $P$, and can be solved in polynominal time \cite{1702662}.

\subsection{Co-observation performance}
Notice that problem \eqref{eq:flow-coverage-problem} is guaranteed to have a solution for $\cK=N_{p}$ where all $\cK$ robots follows the reference trajectory ($f^{k}_{ij}=1, \forall \cI_{v_{i}}=\cI_{v_{j}}=k$).  Additionally, for a fixed number of robots $\cK \geq N_{p}$, it is possible to have a subset of flows $F_{e} \in F$ that is empty, i.e. $f_{ij}=0, \forall (v_{i},v_{j})\in E$, $f\in F_{e}$, these flow will not increase the cost and will not have any impact on the solution.  
The trade-off between the number of surveillance robots and security performance is crucial, with an increase in robots generally enhancing security via cross-trajectory co-observations. However, this improvement can be offset by heightened complexity and coordination challenges. To balance these aspects, the penalty parameter $\rho$ in the cost function \ref{eq:flow-cost} is optimized, achieving an equilibrium between security enhancement and operational efficiency. An iterative approach with $\cK=1$ gradually increases until $F$ contains an empty flow, indicating the point where further robot additions do not improve performance. To control unbounded growth, $\rho$ regulates the second term for a single flow in \eqref{eq:flow-cost} $ \rho \sum_{(ij)\in E} w_{ij} f^k_{ij}$ is always smaller than $1$.

@@

\section{Summary}
In this paper, we provide a method to enhance the security of a multi-robot system without sacrificing the performance. This is done by introducing additional robots to perform cross-trajectory co-observations by traveling between different trajectories. We model the unsecured multi-robot trajectories as a checkpoint graph by identifying checkpoints that requires observation and cross-trajectory paths that can safely access the checkpoints from different trajectories. We have shown that the co-observation planning problem across different trajectories can be transformed into a multi-flow problem on the graph, and that we can find the minimal number of robot needed to finish the co-observation task.
